---
title: "eds213_data_cleaning_assign_GHANADAN"
format: html
editor: visual
---

## Setup & data import

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
# load libraries
library(tidyverse)
```

```{r}
# import data
datadir_processed <- "data/processed/"
data_file <- "snow_cover.csv"
cover_df <- read_csv(file.path(datadir_processed, data_file))

glimpse(cover_df)
```

## Clean 'Water_cover' column

Currently the data type in this column is character, but we want it to be numeric.

```{r}
# inspect values that are NA if values are converted to numeric
cover_df %>% 
  count(Water_cover) %>% 
  filter(is.na(as.numeric(Water_cover)))
```

```{r}
# replace 'n/a' and 'unk' values with NA
cover_df <- cover_df %>% 
  mutate(Water_cover = ifelse(Water_cover %in% c("n/a", "unk"), NA, Water_cover))
```

```{r}
# look at what is going on in df when "." is entered
cover_df %>% 
  filter(Water_cover == ".") %>% 
  View()
```

When "." is entered water cover, it is also the case for all the other covers, so we will just replace these with NA (since there is no supplemental information in the provided metadata).

```{r}
# replace "." values in Water_cover with NAs
cover_df <- cover_df %>% 
  mutate(Water_cover = ifelse(Water_cover == ".", NA, Water_cover))
```

```{r}
# look at what is going on in df when "." is entered
cover_df %>% 
  filter(Water_cover == "-") %>% 
  View()
```

The same thing is going on for "-", so we will replace this with NA as well.

```{r}
# replace "-" values in Water_cover with NAs
cover_df <- cover_df %>% 
  mutate(Water_cover = ifelse(Water_cover == "-", NA, Water_cover))
```

```{r}
# ensure everything was fixed correctly and column is ready to be transformed to numeric data type
cover_df %>% 
  count(Water_cover) %>% 
  filter(is.na(as.numeric(Water_cover)))
```

```{r}
# transform column data type to numeric
cover_df <- cover_df %>% 
  mutate(Water_cover = as.numeric(Water_cover))

glimpse(cover_df)
```

```{r}
# check if all values are between 0 and 100
cover_df %>% 
  filter((Water_cover > 100) | (Water_cover < 0))
```

```{r}
# set Water_cover values for this observation to NA
cover_df <- cover_df %>% 
  mutate(Water_cover = ifelse(Water_cover > 100, NA, Water_cover))
```

## Clean 'Land_cover' column

As with the Water_cover column, the data type in this column should be numeric, but with no processing, it is being read as character.

```{r}
# inspect values that are NA if values are converted to numeric
cover_df %>% 
  count(Land_cover) %>% 
  filter(is.na(as.numeric(Land_cover)))
```

These are the same symbols we saw previously, so we will just go ahead and convert all four of these character strings to NA like we did with Water_cover.

```{r}
# replace '-', '.', 'n/a', and 'unk' values with NA
cover_df <- cover_df %>% 
  mutate(Land_cover = ifelse(Land_cover %in% c("-", ".", "n/a", "unk"), NA, Land_cover))
```

```{r}
# ensure everything was fixed correctly and column is ready to be transformed to numeric data type
cover_df %>% 
  count(Land_cover) %>% 
  filter(is.na(as.numeric(Land_cover)))
```

```{r}
# transform column data type to numeric
cover_df <- cover_df %>% 
  mutate(Land_cover = as.numeric(Land_cover))

glimpse(cover_df)
```

```{r}
# check if all values are between 0 and 100
cover_df %>% 
  filter((Land_cover > 100) | (Land_cover < 0))
```

```{r}
# set Water_cover values for these observation to NA
cover_df <- cover_df %>% 
  mutate(Land_cover = ifelse(Land_cover < 0, NA, Land_cover))
```

## Recompute 'Total_cover' column and infer missing values

The 'Total_cover' column is the 

```{r}
# recompute column by adding snow, water, and land cover percentages
cover_df <- cover_df %>% 
  mutate(Total_cover = Snow_cover + Water_cover + Land_cover)
```

```{r}

```


